### 2015-10-14 latex/CGI_Transvoxels/template_CGIconf.tex

\RequirePackage{fix-cm}

%
%\documentclass{svjour3}                     % onecolumn (standard format)
%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
%\documentclass[smallextended]{svjour3}       % onecolumn (second format)
\documentclass[twocolumn]{cgi2015_latex/svjour3}          % twocolumn
\smartqed  % flush right qed marks, e.g. at end of proof
\usepackage{graphicx}
\usepackage{amsmath, amssymb}
\journalname{CGI2015} % The correct name will be entered by the editor

\graphicspath{{figures/}}

\begin{document}

\title{Real-time visualisation of volumetric data on the GPU}
\subtitle{}
\author{H. Perrier \and J. Dupuy \and J.-C. Iehl \and J.-P. Farrugia}
\institute{H. Perrier \at helene.perrier@liris.cnrs.fr \and J. Dupuy \at jonathan.dupuy@liris.cnrs.fr \and J.-C. Iehl \at jean-claude.iehl@liris.cnrs.fr \and J.-P. Farrugia \at jean-philippe.farrugia@liris.cnrs.fr}
\date{ }% The correct dates will be entered by the editor

\maketitle

\begin{figure}
\centering
\includegraphics[width=0.7\linewidth]{octree}
\includegraphics[width=0.7\linewidth]{mesh_muti_res}
\includegraphics[width=0.7\linewidth]{terrain_potential}
\caption{Our method is based on Lengyel's algorithm \cite{lengyel2010voxel} combined with a linear restricted octree (top) which allows us to create adaptive meshes (middle) from volumetric data in real-time (bottom).}
\end{figure}

\begin{abstract}
We present a tool to visualise volumetric data in real-time by extracting an adaptive triangulated surface from the object.
We extend an existing adaptive marching cubes solution and increase its execution speed by developing a pipeline that is fully data parallel and implemented on the GPU.
This lets us visualise in real-time static scenes as well as fully dynamic data.
Furthermore, we propose a criterion to control the adaptivity of the triangulation that adds control over the projected size of triangles.
This allow us to remove geometric aliasing in the resulting rendering.
\keywords{GPU, Real-time, Level of Details, Volumetric Datasets}
\end{abstract}

\input{introduction}
%p1 : pourquoi volumes cools et important de les visu
%p2 : nos contraintes et pk celles là
%p3 : les methodes actuelles ac leurs objectifs !=
%p4 : ce qu'on fait
%p5 : plan

\input{preliminaries} %previous work
\input{ourmethod} %how/why it works
\input{results} %how awesome it is
\input{conclusion} %how it could be more awesome

\bibliographystyle{alpha}
\bibliography{bibtex}

\appendix

\section{Ensuring a restricted octree}

\paragraph{}
To partition the space, Dupuy \textit{et al}. use the criterion given in equation (\ref{lod_criterion}).
They then test the size of the cell over this criterion.
If equation (\ref{ratio_division}) is verified, it means that the cell will be twice more subdivided than its neighbour.
To maintain a restricted tree, we need to ensure that it never happens.\\
\begin{equation}
\frac{\mathrm{cell\_size}}{2 \cdot s(z)} < 2k
\label{ratio_division}
\end{equation}

\paragraph{}
We illustrate this limit case Figure \ref{proof_constraint}.
Because the only variable is $\alpha$, we need to bound it.
From Thales, we can deduce:
$$
\frac{2k \cdot s(z)}{s(z)} = \frac{z + s(z)}{z}
$$
From this, we can express $z$ in terms of $s(z)$:
$$
z = \frac{s(z)}{2k - 1}
$$
Finally, trigonometry gives us:
$$
tan(\alpha) = 2k - 1
$$
As $\alpha$ is half the horizontal viewing angle, the octree is constraint as long as $\alpha < 2\arctan(2k - 1)$
For $k = 2$, this implies that the fovy must not exceed $142^o$.

\begin{figure} [h]
\centering
\includegraphics[height=0.2\textheight]{proof1}
\caption{If this situation happens, the LoD criterion no longer ensures to maintain a restricted tree.
The parameter $\alpha$ is thus bounded. Over a certain value, a cell can have two level of differences with its neighbours.
This figure is willingly wrongly scaled for readability purposes.}
\label{proof_constraint}
\end{figure}

\end{document}